---
title: "US General Society Survey Analysis"
thanks: "Code and data are available at: [repository](https://github.com/NotSakura/USSocialSurvey.git)"
classoption: abstract
author: "Cristina Burca, Yan Mezhiborsky, Sakura Noskor"
format: pdf
date: today
date-format: long
abstract: |
 This data uses the US 2020 Census results that is in the Harvard Database in order to analyse if there is a correltaion between the election numbers in 2020 and other factors. The factors include immigration status, gender, birth year, and results from the 2016 election. It was found that there does exist a correlation between the outcomes of the 2020 election and the above variables. These are important results as the US is a country that si considered a hub, due to all the external relations they have with multiple countries. These results will also help predict who may win the 2024 elections. 
include-in-header: 
  text:
    \renewcommand{\abstractname}{Abstract}
bibliography: references.bib
margins:
  top: 0.25in
  bottom: 0.25in
  left: 0.5in
  right: 0.5in
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(patchwork)
library(gridExtra)
library(rstanarm)
library(modelsummary)
library(dplyr)
library(lubridate)
library(kableExtra)
library(knitr)
```

```{r}
#| include: false
#| warning: false
#| message: false

library(dataverse)
library(tidyverse)

ces2020 <-
  get_dataframe_by_name(
    filename = "CES20_Common_OUTPUT_vv.csv",
    dataset = "10.7910/DVN/E9N6PH",
    server = "dataverse.harvard.edu",
    .f = read_csv
  ) |>
  select(votereg, CC20_410, immstat, gender, birthyr, presvote16post)

write_csv(ces2020, "ces2020.csv")



```

```{r}
#| include: false
#| warning: false
#| message: false


ces2020 <-
  read_csv(
    "ces2020.csv",
    col_types =
      cols(
        "votereg" = col_integer(),
        "CC20_410" = col_integer(),
        "immstat" = col_integer(), 
        "gender" = col_integer(), 
        "birthyr" = col_integer(),  
        "presvote16post" = col_integer(), 
      )
  )

```

```{r}
#| include: false
#| warning: false
#| message: false

ces2020 <-
  ces2020 |>
  filter(votereg == 1,
         CC20_410 %in% c(1, 2)) |>
  mutate(
    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
    voted_for = as_factor(voted_for),
    gender = if_else(gender == 1, "Male", "Female"),
    immstat = case_when(
      immstat == 1 ~ "immigrant and citizen",
      immstat == 2 ~ "immigrant not citizen",
      immstat == 3 ~ "born in US, but parent(s) immigrant",
      immstat == 4 ~ "parent and I born in US but grandparent(s) immigrant",
      immstat == 5 ~ "all born in US",
    ),
    immstat = factor(immstat, levels = c( "immigrant and citizen","immigrant not citizen", "born in US, but parent(s) immigrant",     "parent and I born in US but grandparent(s) immigrant", "all born in US")
    ),
    presvote16post = case_when(
      presvote16post == 1 ~ "Donald Trump",
      presvote16post == 2 ~ "Hillary Clinton",
      presvote16post == 3 ~ "Gary Johnson",
      presvote16post == 4 ~ "Jill Stein",
      presvote16post == 5 ~ "Evan McMullin",
      presvote16post == 6 ~ "Other", 
      presvote16post == 7 ~ "Did not vote"
    ),
    presvote16post = factor(presvote16post, levels = c(
        "Hillary Clinton","Donald Trump","Gary Johnson", "Jill Stein", "Evan McMullin", "Other", "Did not vote")
    )
  ) |>
  select(voted_for, votereg, immstat, gender, birthyr, presvote16post)

```


# Introduction

The United States is one of the leading countries in export, imports and almost everything economic and socially related. The US contributes to the worlds economy by 20% despite the fact that they contain 5% of the populations[@USTradeRep]. This makes the US very relevant to not only national news but also international news. This is why the US presidential elections are broadcaster to worldwide ever election term. The results of the election not only affect American citizens but also external affairs related to the country. 

The United States of America is a democratic government which means that they hold election every 4 years[@usc_presidential_election]. The 2 parties that historically run against each other with the most votes are the Republican party and the Democratic Party. The republicans are often associated with conservative beliefs and values such as views opposing abortion and privatization to save their economy. The Democrats, on the other hand, are often associated with liberal views such as social welfare programs and higher taxes to support the government aids provided to citizens[@britannica_dem_vs_rep]. There are many other parties such as Libertarian Party, Green Party, Constitution Party and other independent candidates but because majority of the votes goes to the 2 parties the others are often over looked. The legal voting age is 18 in the states and you must be a registered voter in order to take part of the election which means non-citizens are not taken into account. Students are not as well. Also there seems to be a 66% voter turnout which means the remaining 34% decided not to vote [@DeSilver_2022]. The 2024 elections is also coming up this year meaning that analyzing this results may help us predict what the outcome of the election might be. 

This research examines the voting patterns in the 2016 and 2020 US Presidential elections. We will be looking at data collected by the Cooperative Election Study and accessed through the Harvard University Database [@dataverse]. The analysis is based on a representative sample of 61,000 American adults, which provides detailed information about each individuals gender, birth year, race, registered state, employment, education loans, immigration status, dual-citizenship, religion, and 2016 and 2020 Presidential vote. The goal of this study is to use relevant variables from the electoral data to investigate patterns, trend and predictions regrading American electoral preferences from 2016 and 2020.

# Data {#sec-data}


## Data Used
This paper was modeled with the help of R [@citeR] along with other useful packages like tidyverse [@tidy] (which includes graphing functions like ggplot2), patchwork [@patch]. There are parts of the code which were guided by Rohan Alexander's Telling Stories with fire [@rohan] chapter 13 section 13.2.2. The data was used from the Harvard database [@dataverse]

## Variables inspected
Starting off, we examine the columns 'votereg' and 'voted_for'. They represent the number of persons that registered to vote and which candidate they voted for in 2020, respectively. We filtered out the rows with a 'votereg' value of 2, which indicated unregistered voters, to focus exclusively on individuals who were registered to vote. 
We then focused on the 'presvote16post' variable, which reveals the candidates Americans voted for in the 2016 United States Presidential Election. This is an important variable as it enables us to assess whether American citizens were satisfied with the service that the previous government provided. Next we look at 'gender' as well as 'employment'. Both 'gender' and 'employment' shows us if there is a correlation between certain parties views versus the demographic they represent. 'Gender' contains 2 values (male and female) while, employment has 9 values; full time worker, part time worker, laid off, unemployed, retired, permanently disabled individual, Homemaker, Student or Other. We also explore the variable 'immstat' which represents the immigration status of the of individual represented by one of the following: immigrant and citizen, immigrant not citizen, born in US, but parent(s) immigrant, parent and I born in US but grandparent(s) immigrant, or all born in US. 


## The Destination to Reach with the Data
There could have been many other similar data sets that could have been used for this project for example we could have chosen to look at the census and election data for Canada. However, our group decided that because part of the analysis was done in @rohan, there were still many other variables that we could explore as we dive further into the 2020 presidential election and try to interpret if there are any correlations between the variables and the result. Our team found it interesting to see all the variables that were collected by the US government and the correlations we saw during the analysis process; where there most definitely was a positive correlation between each variable and the outcome of the votes. Although we are analyzing the 2020 election that has already taken place, the analysis we do in the later sections are believed to apply to the 2024 elections happening this year. This is enough reason for us and the reader to dive into the patterns that exist with this large data set.


# Models/Results

## Aging and Its Impact on Voting

In these following models we will of our research paper, we will look at the relationship between the demographic of age within the population and their voting patterns.The focus will primarily be on understanding how different age groups align themselves politically. By investigating the percentage of voters within each age bracket who support particular candidates, we aim to uncover nuanced insights into the demographic underpinnings of political support.Through the integration of descriptive statistics, analysis, and visual data representation we will offer a comprehensive overview on age base voting patterns. This will not only provide empirical evidence on the voting preferences of different ages but also help for a deeper understanding of the potential motivations behind their voting choices.

### Aging Voters (2016)

In the following graphs we will be looking at the percentage of voters from each birth year that voted for Hillary Clinton, Donald Trump,and other candidates in the 2016 US Presidential election, we will also take a look at those who did not vote.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-hillary2016
#| fig-cap: "Comparing Percentage of Votes For Hillary Clinton and Year of Birth (2016)"


ces2020_votes <- ces2020 %>%
  filter(presvote16post %in% c("Hillary Clinton", "Donald Trump", "Other", "Did not vote"))

# Calculate percentages
percentage_data <- ces2020_votes %>%
  group_by(birthyr) %>%
  summarise(
    HillaryClinton = sum(presvote16post == "Hillary Clinton") / n() * 100,
    DonaldTrump = sum(presvote16post == "Donald Trump") / n() * 100,
    Other = sum(presvote16post == "Other") / n() * 100,
    DidNotVote = sum(presvote16post == "Did not vote") / n() * 100
  ) %>%
  ungroup()

# Reshape data to long format
percentage_data_long <- pivot_longer(percentage_data, cols = -birthyr, names_to = "Candidate", values_to = "Percentage")

#R value
hillary_data <- percentage_data_long %>% filter(Candidate == "HillaryClinton")

R_hillary <- cor(hillary_data$birthyr, hillary_data$Percentage)

plot_hillary <- ggplot(percentage_data_long %>% filter(Candidate == "HillaryClinton"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Percentage for Hillary Clinton", y = "Percentage (%)", x = "Birth Year") +
  theme_minimal()
plot_hillary

```

In @fig-hillary2016 we can see that the younger the voters, the smaller percentage of their birth year seem to have voted for Hillary Clinton in the 2016 presidential election. This was rather surprising for us, as we expected young voters more likely to vote for Hillary Clinton and the Democratic Party, as younger people are often more progressive than their older counterparts. @fig-hillary shows us that there seems to be a significant correlation between age, birth year, and what percent of that birth year voted for Hillary with a correlation coefficient of $R = -0.94$, which is a rather strong negative relationship between our two variables.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-trump2016
#| fig-cap: "Comparing Percentage of Votes For Donald Trump and Year of Birth (2016)"

trump_data <- percentage_data_long %>% filter(Candidate == "DonaldTrump")

R_trump <- cor(trump_data$birthyr, trump_data$Percentage)

#Trum Plot
plot_trump <- ggplot(percentage_data_long %>% filter(Candidate == "DonaldTrump"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Percentage for Donald Trump", y = "Percentage (%)", x = "Birth Year") +
  theme_minimal()
plot_trump
```

While we look at the graph for Donald Trump in the 2016 election at first glance it does seem that the younger you are the more likey you were to have voted for trump in the 2016 election, but if we dig a bit deeper we can see that there seem to be several outliers on this graph, and the data points seem to somewhat be bouncing around a bit. This is confirmed when we look at our correlation coefficient $R = 0.12$, so we can come to a conclsion that voting for Donald Trump in 2016 has a relatively weak positive correlation. 

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-other2016
#| fig-cap: "Comparing Percentage of Votes For Other Caniddates and Year of Birth (2016)"

other_data <- percentage_data_long %>% filter(Candidate == "Other")

R_other <- cor(other_data$birthyr, other_data$Percentage)

plot_other <- ggplot(percentage_data_long %>% filter(Candidate == "Other"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "green") +
  labs(title = "Percentage for Other", x = "Birth Year", y = "Percentage (%)") +
  theme_minimal()
plot_other
```

@fig-other2016 does not seem to give us to much information about the age of those who voted for other candidates in the 2016 presidential elections, such as Gary Johnson and Jill Stein. While all these candidates received a combined less than 5% of the total vote, and less than 5% of the share of the vote from each birth year, we believe that transparency is incredibly important when doing data heavy research and while this may not seem as important as the graphs for Hillary Clinton and Donald Trump, this graph absolutely has a place in this paper.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-dnv2016
#| fig-cap: "Comparing Percentage of Non-Voters and Year of Birth (2016)"

dnv_data <- percentage_data_long %>% filter(Candidate == "Did not vote")

R_dnv <- cor(dnv_data$birthyr, dnv_data$Percentage)

#plot did not vote
plot_did_not_vote <- ggplot(percentage_data_long %>% filter(Candidate == "DidNotVote"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "orange") +
  labs(title = "Percentage for Did Not Vote", x = "Birth Year", y = "Percentage (%)") +
  theme_minimal()
plot_did_not_vote
```

Here we will look at the members of the population that did not vote. @fig-dnv2016 shows us the steady increase of those citizens who did not vote in the 2016 presidential elections. We can see that they younger generation seems to have not voted for any candidate in the 2016 elections.  This can be due to a variety of reasons, such as disengagement with the political process, dissatisfaction with the candidates presented, or a feeling that their vote may not make a significant difference in the outcome. Other contributing factors could include obstacles to voting access, such as registration issues, polling place locations, or work and school schedules that conflict with voting hours.

As we further dissect these trends, it's essential to understand the broader societal and cultural influences that might be contributing to this detachment.

### Aging Voters (2020)

The data provided for the 2020 elections was slightly different the the data from the 2016 elections, the main difference was that, there was no one who voted for another candidate, other than Joe Biden and Donald Trump from the Democratic and Republican parties respectively. There were also no "Did not vote" inputs like there were in the data from the 2016 election between Donald and Hillary.


```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-biden2020
#| fig-cap: "Comparing Percentage of Votes For Joe Biden and Year of Birth (2020)"

ces2020_votes <- ces2020 %>%
  filter(voted_for %in% c("Biden", "Trump"))

# Calculate percentages for all categories
percentage_data_2 <- ces2020_votes %>%
  group_by(birthyr) %>%
  summarise(
    Biden = sum(voted_for == "Biden") / n() * 100,
    Trump = sum(voted_for == "Trump") / n() * 100
  ) %>%
  ungroup()

# Reshape data to long format
percentage_data_2_long <- pivot_longer(percentage_data_2, cols = -birthyr, names_to = "Candidate", values_to = "Percentage")

# Plot for Joe Biden
plot_biden <- ggplot(percentage_data_2_long %>% filter(Candidate == "Biden"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Percentage for Joe Biden", y = "Percentage (%)", x = "Birth Year") +
  theme_minimal()

plot_biden

#Rvalue Biden
biden_data <- percentage_data_long %>% filter(Candidate == "Biden")

R_biden <- cor(biden_data$birthyr, biden_data$Percentage)


```

In @fig-biden2020 we see a complete opposite graph compared to Hillary's from 2016 where her's seemed to be more popular among older voters where as Biden seems to be much more popular among younger voters. So much so that voters born after 1980 had at least 60% of their age groups voting for Biden and those born after 1990 had at least 70% of their age groups voting for Biden. This is a much less surprising Graph than Hillary's, as younger voters seem to be more likely to vote for more progressive parties and candidates.

We can see that for the most part the graph seems to be increasing at nearly every point except for a dip starting around 1950 and does not start to pick up until 1960, which seems somewhat odd until we recall that those are the prime birth year of the baby boomer generation.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-trump2020
#| fig-cap: "Comparing Percentage of Votes For Donald Trump Year of Birth (2020)"

# Plot for Donald Trump
plot_trump_2 <- ggplot(percentage_data_2_long %>% filter(Candidate == "Trump"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Percentage for Donald Trump", y = "Percentage (%)", x = "Birth Year") +
  theme_minimal()

plot_trump_2

#rvalue Trump 2020

trump_2_data <- percentage_data_long %>% filter(Candidate == "Trump")

R_trump_2 <- cor(trump_2_data$birthyr, trump_2_data$Percentage)
```


Lastly, we will look at @fig-trump2020 which is essentially the inverse of the previous graph as there were no other possible data points other than "Biden" and "Trump". While this graph may just be a reflection across the 50% line, it is still helpful so that wer may look at it from a different perspective. Just as in @fig-biden2020, @fig-trump2020 has a small peak after the year 1950. It is also important to note that even though there is a peak after the year 1950, all of the ages that had more than a 50% vote share for trump all also came before the year 1950.


## Model 2
In this model, we conduct an examination of the relationship between voters' birth year and their gender for the 2016 and 2020 Presidential elections. This analysis is visualized by plotting a histogram that separate female voters on the left and male voters on the right, with voters' birth years measured along the x-axis, which ranges from 1925 to 2002. The y-axis quantifies the voter turnout for the year. For clarity ad symbolic representation, the colour blue was chosen to represent the Democratic candidates --Hillary Clinton for the 2016 election, and Joe Biden for the 2020 election, while red was chosen to the Represent the Republican candidate, Donald Trump, who sought the presidency in both terms. @fig-genderyr2016 present the distribution of votes in 2016, and @fig-genderyr2020 presents the data from the 2020 election. 


```{r, fig.height=6, fig.width=11}
#| echo: false
#| warning: false
#| message: false
#| label: fig-genderyr2016
#| fig-cap: "Logistic regression of 2016 US Presidential votes comparing parameters of gender and immigration status"

my_palette <- c("Hillary Clinton" = "skyblue3", "Donald Trump" = "indianred3")
my_palette2 <- c("Biden" = "skyblue3", "Trump" = "indianred3")

ces2020_votes %>% 
  ggplot(aes(x= birthyr, fill=presvote16post)) +
  geom_histogram(binwidth = 2, position = "dodge", alpha = 0.8) + 
  facet_wrap(facets = vars(gender), scales = "free_x") +
  theme_minimal() +
  scale_fill_manual(values = my_palette) + 
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(x ="Birth year",
      y = "Count of votes",
      fill = "Vote in 2016")
```
We notice in @fig-genderyr2016 the graph displays two distinct high points for the number of female Republican voters, with the peak around 1960 being the most pronounced, followed by another around 1980. There are similar peaks in the graph for female Democratic voters, but with the overall count being considerably lower. We also notice that the Democratic party received slightly more votes from the older demographic, where as the younger demographic greatly preferred the Republican party. In comparison, the graph displaying the male votes has a more balanced distribution, and similar to the female voters, the older and younger demographic preferred the Democratic Party and the Republican Party, respectively. It is worthy to mention that there is a higher count of women than men.



```{r, fig.height=6, fig.width=11}
#| echo: false
#| warning: false
#| message: false
#| label: fig-genderyr2020
#| fig-cap: 'Logistic regression of 2020 US Presidential votes comparing parameters of gender and immigration status'

#plot for 2020 votes by 
ces2020_votes %>% 
  ggplot(aes(x= birthyr, fill=voted_for)) +
  geom_histogram(binwidth = 2, position = "dodge", alpha = 0.8) + 
  facet_wrap(facets = vars(gender), scales = "free_x") +
  theme_minimal() +
  scale_fill_manual(values = my_palette2) + 
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(x ="Birth Year",
      y = "Count of votes",
      fill = "Vote in 2020")

```
We notice in @fig-genderyr2020 that the graphs bear a strong resemblance to the distributions of the 2016 votes depicted in @fig-genderyr2016, but with the parties reversed on the graph. We are interested in the ratios of these graphs.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-ratio2016

ces2020_votes_2016 <- ces2020_votes %>%
  mutate(birth_decade = (birthyr %/% 10) * 10,
    trump_vote = ifelse(presvote16post == "Donald Trump", 1, 0),
    biden_vote = ifelse(presvote16post == "Hillary Clinton", 1, 0))

ces2020_votes_2016 <- ces2020_votes_2016 %>%
  filter(!(birth_decade %in% c(1920, 2000)))

votes_count <- ces2020_votes_2016 %>%
  group_by(birth_decade, gender) %>%
  summarise(biden_vote_total = sum(biden_vote), trump_vote_total = sum(trump_vote), .groups = "drop")

votes_count <- votes_count %>%
  mutate(vote_ratio = biden_vote_total / trump_vote_total)

votes_count_female <- votes_count %>% 
  filter(gender == "Female")
votes_count_male <- votes_count %>% 
  filter(gender == "Male")

votes_count_female <- select(votes_count_female, -gender)
votes_count_male <- select(votes_count_male, -gender)

colnames(votes_count_female) <- c("Birth Decade", "Total Clinton Votes", "Total Trump Votes", "Ratio")
colnames(votes_count_male) <- c("Birth Decade", "Total Clinton Votes", "Total Trump Votes", "Ratio")

votes_2020_f <- kable(votes_count_female, 
      caption = "Ratio of 2016 Votes for Clinton to Trump by Birth Decade (Females)",
      digits = 2)

votes_2020_m <- kable(votes_count_male, 
      caption = "Ratio of 2016 Votes for Clinton to Trump by Birth Decade (Males)",
      digits = 2)

kables(list(votes_2020_f, votes_2020_m))
#If the ratio is greater than 1, it means that within the group, Biden has more votes than Trump
#if the ratio is less than 1, Trump has more votes within the group
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-ratio2020

ces2020_votes_2020 <- ces2020_votes %>%
  mutate(birth_decade = (birthyr %/% 10) * 10,
    trump_vote = ifelse(voted_for == "Trump", 1, 0),
    biden_vote = ifelse(voted_for == "Biden", 1, 0))

ces2020_votes_2020 <- ces2020_votes_2020 %>%
  filter(!(birth_decade %in% c(1920, 2000)))

votes_count <- ces2020_votes_2020 %>%
  group_by(birth_decade, gender) %>%
  summarise(biden_vote_total = sum(biden_vote), trump_vote_total = sum(trump_vote), .groups = "drop")

votes_count <- votes_count %>%
  mutate(vote_ratio = biden_vote_total / trump_vote_total)

votes_count_female <- votes_count %>% 
  filter(gender == "Female")
votes_count_male <- votes_count %>% 
  filter(gender == "Male")

votes_count_female <- select(votes_count_female, -gender)
votes_count_male <- select(votes_count_male, -gender)

colnames(votes_count_female) <- c("Birth Decade", "Total Biden Votes", "Total Trump Votes", "Ratio")
colnames(votes_count_male) <- c("Birth Decade", "Total Biden Votes", "Total Trump Votes", "Ratio")

votes_2020_f <- kable(votes_count_female, 
      caption = "Ratio of 2020 Votes for Biden to Trump by Birth Decade (Females)",
      digits = 2)

votes_2020_m <- kable(votes_count_male, 
      caption = "Ratio of 2020 Votes for Biden to Trump by Birth Decade (Males)",
      digits = 2)

kables(list(votes_2020_f, votes_2020_m))
#If the ratio is greater than 1, it means that within the group, Biden has more votes than Trump
#if the ratio is less than 1, Trump has more votes within the group
```

We analyze the ratio of Democratic to Republic votes in 2016 in @tbl-ratio2016 and in 2020 in @tbl-ratio2020. We divide respondents by birth decade, and we list the total number of votes for the Democratic Party and thee Republican Party, then list the ratio of the two. The decades 1920 and 2000 were omitted since the number of respondents in those Birth Decades were very low. The value of the ratio in centered at 1, where if the values is greater than 1, then within that group, the Democratic Party has more votes than the Republican Party. Where the values are less than 1, the Republican Party has more votes. When the ratio is around 1, the number of votes is approximately equal. 

We notice that in 2016, the ratio of votes decreases as birth decade increases for both women and men voters. The opposite happens in 2020, where the ratio of votes increases as birth decade increases. 

## Model 3

We have modeled the following logistic regression in the graphs:

$$
\begin{aligned} 
y_i|\pi_i  &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{gender}_i + \beta_2\times \mbox{immigration status}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5)
\end{aligned}
$$

```{r, fig.align='right', fig.height=5, fig.width=11}
#| echo: false
#| warning: false
#| message: false


#plot for 2016 votes
ces2020_votes %>% 
  ggplot(aes(x= immstat, fill=presvote16post)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(gender), scales = "free_x") +
  theme_minimal() +
  coord_flip() +
  scale_fill_manual(values = my_palette) +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(x ="Immigration status",
      y = "Count of votes",
      fill = "Vote in 2016")

#plot for 2020 votes
ces2020_votes %>% 
  ggplot(aes(x= immstat, fill=voted_for)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(gender), scales = "free_x") +
  theme_minimal() +
  coord_flip() +
  scale_fill_manual(values = my_palette2) + 
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(x ="Immigration status",
      y = "Count of votes",
      fill = "Vote in 2020")


#trump republican hillary and bidne were democrat 
#male born in the US are a 50/50, everyone else changed minds
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

set.seed(302)

ces2020_reduced <- 
  ces2020_votes |> 
  slice_sample(n = 1000)

political_preferences <-
  stan_glm(
    voted_for ~ gender + immstat,
    data = ces2020_reduced,
    family = binomial(link = "logit"),
    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
    prior_intercept = 
      normal(location = 0, scale = 2.5, autoscale = TRUE),
    seed = 302
  )

saveRDS(
  political_preferences,
  file = "../models/political_preferences.rds"
)

political_preferences <-
  readRDS(file = "../models/political_preferences.rds")

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of presidential election in 2020 based on gender and immigration status"
#| warning: false
modelsummary(
  list(
    "Support Biden" = political_preferences
  ),
  statistic = "mad"
  )
```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

  It is near impossible to have data sets that are completely error-free and the same is true for this data-set. The data-set we used as analysis does not take into account the opinions of non-citizens and who they would have voted for. A lot of marginalized opinions are being left out when we clean the data and we only look for people that could vote. For example refugees may have a different opinion on who got elected in the 2020 presidential candidate as they come from a different aspect of life and have different priorities that need to be addressed such as housing and or health care. 
  Another weakness in our data-set is what occurs with a lot of data-sets, which is missing values. This data set also contained them and those needed to be cleaned so that our analysis would not sway. Speaking of missing values, insignificant values were ignored in some of the models as well. For, example some of the presidential candidates got so little votes that including them would have actually hindered our ability to see the trends so we had to group them in others. This was beneficial for this paper but  if we do indeed continue to analyse this paper for other purposes, this might cause problems. 
  
  In terms of next steps for this paper, we look forward to predict the 2024 election results by comparing the what the general priorities of the population is. We could look at population by immigration status, gender, age, or their previous 2020 voting results.  If we are able to correctly predict the results based on the model continuously, it would help understand the priorities of american citizens; knowledge that would in turn help us help them. 

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(political_preferences) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(political_preferences) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()

```

## Diagnostics

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2


plot(political_preferences, "trace")

plot(political_preferences, "rhat")

```

\newpage

# References
