---
title: "US General Society Survey Analysis"
thanks: "Code and data are available at: [repository](https://github.com/NotSakura/USSocialSurvey.git)"
classoption: abstract
author: "Cristina Burca, Yan Mezhiborsky, Sakura Noskor"
format: pdf
date: today
date-format: long
abstract: |
 Some abstact
include-in-header: 
  text:
    \renewcommand{\abstractname}{Abstract}
bibliography: references.bib
margins:
  top: 0.25in
  bottom: 0.25in
  left: 0.5in
  right: 0.5in
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(patchwork)
```



```{r}
#| include: false
#| warning: false
#| message: false

library(dataverse)
library(tidyverse)

ces2020 <-
  get_dataframe_by_name(
    filename = "CES20_Common_OUTPUT_vv.csv",
    dataset = "10.7910/DVN/E9N6PH",
    server = "dataverse.harvard.edu",
    .f = read_csv
  ) |>
  select(votereg, CC20_410, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)

write_csv(ces2020, "ces2020.csv")



```

```{r}
#| include: false
#| warning: false
#| message: false


ces2020 <-
  read_csv(
    "ces2020.csv",
    col_types =
      cols(
        "votereg" = col_integer(),
        "CC20_410" = col_integer(),
        "employ" = col_integer(),
        "dualcit" = col_integer(), 
        "gender" = col_integer(), 
        "birthyr" = col_integer(), 
        "race" = col_integer(), 
        "inputstate" = col_integer(), 
        "presvote16post" = col_integer(), 
        "religpew" = col_integer(), 
        "edloan" = col_integer()
      )
  )

```



```{r}
#| include: false
#| warning: false
#| message: false

ces2020 <-
  ces2020 |>
  filter(votereg == 1,
         CC20_410 %in% c(1, 2)) |>
  mutate(
    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
    voted_for = as_factor(voted_for),
    gender = if_else(gender == 1, "Male", "Female"),
    employ = case_when(
      employ == 1 ~ "Full-time",
      employ == 2 ~ "Part-time",
      employ == 3 ~ "Laid off",
      employ == 4 ~ "Unemployed",
      employ == 5 ~ "Retired",
      employ == 6 ~ "Permannetly disabled", 
      employ == 7 ~ "Homemaker", 
      employ == 8 ~ "Student", 
      employ == 9 ~ "Other"
    ),
    employ = factor(
      employ,
      levels = c(
        "Full-time",
        "Part-time",
        "laid off",
        "unemployed",
        "retired",
        "permannetly disabled", 
        "Homemaker", 
        "Student", "Other"
      )
    ), 
    immstat = case_when(
      immstat == 1 ~ "immigrant and citizen",
      immstat == 2 ~ "immigrant not citizen",
      immstat == 3 ~ "born in US, but parent(s) immigrant",
      immstat == 4 ~ "parent and I born in US but grandparent(s) immigrant",
      immstat == 5 ~ "all born in US",
    ),
    immstat = factor(immstat, levels = c( "immigrant and citizen","immigrant not citizen", "born in US, but parent(s) immigrant",     "parent and I born in US but grandparent(s) immigrant", "all born in US")
    ),
    dualcit = if_else(dualcit == 1, "Yes", "No"),
    race = case_when(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Hispanic",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 6 ~ "Middle Eastern", 
      race == 7 ~ "Two or more races", 
      race == 8 ~ "Other"
    ),
    race = factor(race, levels = c(
        "White","Black","Hispanic", "Asian", "Native American", "Middle Eastern",  "Two or more races", "Other")
    ), 
    presvote16post = case_when(
      presvote16post == 1 ~ "Donald Trump",
      presvote16post == 2 ~ "Hillary Clinton",
      presvote16post == 3 ~ "Gary Johnson",
      presvote16post == 4 ~ "Jill Stein",
      presvote16post == 5 ~ "Evan McMullin",
      presvote16post == 6 ~ "Other", 
      presvote16post == 7 ~ "Did not vote"
    ),
    presvote16post = factor(presvote16post, levels = c(
        "Hillary Clinton","Donald Trump","Gary Johnson", "Jill Stein", "Evan McMullin", "Other", "Did not vote")
    ),
    religpew = case_when(
      religpew == 1 ~ "Protestant",
      religpew == 2 ~ "Roman Catholic",
      religpew == 3 ~ "Mormon",
      religpew == 4 ~ "Eastern or Greek Orthodox",
      religpew == 5 ~ "Jewish",
      religpew == 6 ~ "Muslim", 
      religpew == 7 ~ "Buddhist", 
      religpew == 8 ~ "Hindu", 
      religpew == 9 ~ "Atheist", 
      religpew == 10 ~ "Agnostic",
      religpew == 11 ~ "Nothing in Particular",
      religpew == 12 ~ "Something else"
    ),
    religpew = factor(
      religpew,
      levels = c(
        "Protestant",
        "Roman Catholic",
        "Mormon",
        "Eastern or Greek Orthodox",
        "Jewish",
        "Muslim", 
        "Buddhist", 
        "Hindu", "Atheist", "Agnostic", "Nothing in Particular", "Something else"
      )
    ), 
    edloan = case_when(
      edloan == 1 ~ "Yes",
      edloan == 2 ~ "No",
      edloan == 3 ~ "Other",
      edloan == 4 ~ "Don't know",
      edloan == 5 ~ "None of these",
    ),
    edloan = factor(edloan, levels = c( "Yes","No", "Other", "Don't Know", "None of these")
    ) #########ADD input state it if you want to (sakura)
  ) |>
  select(votereg, employ, immstat, dualcit, gender, birthyr, race, inputstate, presvote16post, religpew, edloan)

```


# Introduction




# Data {#sec-data}

- included variables of gender, birth year, race, registered state, employment, education loans, immigration status, dual-citizenship, religion, and 2016 US vote

## Data Used
This paper was modeled with the help of R [@citeR] along with other useful packages like tidyverse [@tidy] (which includes graphing functions like ggplot2), patchwork [@patch]. There are parts of the code which were guided by Rohan Alexander's Telling Stories with fire [@rohan] chapter 13 section 13.2.2.

## Variables inspected
Starting off we look at the column votereg and CS20_410. They represent the number of people that registered to vote and who they voted for, respectively. The columns where the votereg was 2, meant that the person was not registered to vote so we essentially cleaned up those values to only look at people who did vote. The next variable we isolated was 2016pres which tells us, who Americans voted for during the 2016 United States Presidential Election. This is an important variable as we will be able to see if the citizens of america were satisfied with the service that the previous government provided. Then we look at gender as well as employment. Both gender and employment shows us if there is a correlation between a certain parties views vs the demographic they are a part of and represent. Gender contains 2 values (male and female) while, employment has 9 values; full time worker, part time worker, laid off, unemployed, retired, permanently disabled individual, Homemaker, Student or Other. 


## Data Explained
There could have been many other similar datasets that could have been used for this project for example we could have chosen to look at the census and election data for Canada. However, our group decided that because part of the analysis was done in @rohan, there were still many other variables that we could explore as we dive furthur into the 2020 presidential election and try to intrepret if there are any correlations between the variables and the result. Our team found it interesting to see all the variables that were collected by the US government and the correlations we saw during the analysis process; where there most definitely was a positive correlation between each variable and the outcome of the votes. Although we are analyzing the 2020 election that has already taken place, the analysis we do in the later sections are believed to apply to the 2024 elections happening this year. This is enough reason for us and the reader to dive into the patterns that exist with this large dataset.


# Model

The variable 'presvote16post' captures the 2016 US Presidential votes of individuals in the United States. Due to the minimal votes for Gary Johnson, Jill Stein, Evan McMullin, and all other candidates any votes not attributed to 'Hillary Clinton' or 'Donald Trump' were categorized as 'Other'.

We have modeled the following logisitc regression in @fig-votes:



```{r, fig.height=4, fig.width=8}
#| echo: false
#| warning: false
#| message: false
#| label: fig-votes
#| fig-cap: 'Logistic regression of 2016 US Presidential votes comparing parameters of gender and employment status'

#logisitc regression for binary outcome: YES OR NO, THIS OR THAT

#choose values "Hillary Clinton" and "Donald Trump", all other values as "Other"
ces2020_votes <- ces2020 %>%
  filter(presvote16post != 'Did not vote') %>%
  mutate(presvote16post = case_when(
    presvote16post == "Hillary Clinton" ~ "Hillary Clinton",
    presvote16post == "Donald Trump" ~ "Donald Trump",
    presvote16post == "Gary Johnson" ~ "Other",
    presvote16post == "Jill Stein" ~ "Other",
    presvote16post == "Evan McMullin" ~ "Other",
    presvote16post == "Other" ~ "Other"
  ))

#plot 2016 votes against employment status and gender
#SO MANY NA VALUES???
ces2020_votes %>%
  ggplot(aes(x= employ, fill=presvote16post)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(gender)) +
  theme_minimal() +
  coord_flip() +
  labs(x ="Employment status",
      y = "Count of votes",
      fill = "Vote in 2016")
  
```

```{r, fig.height=4, fig.width=8}
#| echo: false
#| warning: false
#| message: false

ces2020_votes <- ces2020_votes %>%
  filter(!is.na(dualcit))

ces2020_votes %>% 
  ggplot(aes(x= immstat, fill=presvote16post)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(dualcit), scales = "free_x") +
  theme_minimal() +
  coord_flip() +
  labs(x ="Employment status",
      y = "Count of votes",
      fill = "Vote in 2016")

```

```{r, fig.height=4, fig.width=8}
#| echo: false
#| warning: false
#| message: false


ces2020_votes %>% 
  ggplot(aes(x= race, fill=presvote16post)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(gender), scales = "free_x") +
  theme_minimal() +
  coord_flip() +
  labs(x ="Employment status",
      y = "Count of votes",
      fill = "Vote in 2016")

```

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results





# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics



```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2


```



\newpage


# References


